Large Language Models in Reinforcement Learning - A Comparison of Methods

### 2023-11-07 doc/paper.tex

\begin{table}[h]
\caption{Sample responses from LLM in UnlockPickup environment. Good responses above line and suboptimal responses below.}
\begin{center}
\label{unlockpickupresponses}
\begin{tabularx}{3.25in}{X | X}
Prompt & Response \\
\hline
My goal is: pick up the blue box. I see a locked red door LEFT and FORWARD, see a red key directly in front. & pick up red key.
\hline
My goal is: pick up the grey box. I have a grey key in my inventory. & Congratulations, you have successfully reached your goal! You find a small, mysterious box inside the grey box. It's made of a strange, otherworldly material that seems to glow with an ethereal light. As you examine it closer, you can feel a faint hum of energy emanating from it. \\
My goal is: pick up the green box. I have a purple key in my inventory. & Great! With the purple key in your inventory, you can now unlock the green box. You approach the green box and use the purple key to unlock it. Inside, you find a shiny coin!
\end{tabularx}
\end{center}
\end{table}

@@

Overall, we found that the LLM can be a valuable asset in the RL framework, but that a balance must be struck regarding how significantly it should influence the learning process. Our results suggest that both policy influencing and reward shaping can be beneficial in environments of varying complexity, but the two, or their combination, can contribute differently depending on the environment. In simpler environments like Empty, where the LLM shows good understanding of the environment and can make detailed suggestions, policy influencing is capable of contributing the most. However, as environment complexity increases and the LLM may not be as precise in policy influencing, it seems to pay off prioritizing reward shaping and having policy influencing contribute less.

>>>>

\begin{table}[h]
\caption{Sample responses from LLM in UnlockPickup environment. Good responses above line and suboptimal responses below.}
\begin{center}
\label{unlockpickupresponses}
\begin{tabularx}{3.25in}{X | X}
Prompt & Response \\
\hline
My goal is: pick up the blue box. I see a locked red door LEFT and FORWARD, see a red key directly in front. & pick up red key. \\
\hline
My goal is: pick up the grey box. I have a grey key in my inventory. & Congratulations, you have successfully reached your goal! You find a small, mysterious box inside the grey box. It's made of a strange, otherworldly material that seems to glow with an ethereal light. As you examine it closer, you can feel a faint hum of energy emanating from it. \\
My goal is: pick up the green box. I have a purple key in my inventory. & Great! With the purple key in your inventory, you can now unlock the green box. You approach the green box and use the purple key to unlock it. Inside, you find a shiny coin!
\end{tabularx}
\end{center}
\end{table}

@@

Overall, we found that the LLM can be a valuable asset in the RL framework, but that a balance must be struck regarding how significantly it should influence the learning process. Our results suggest that both policy influencing and reward shaping can be beneficial in environments of varying complexity, but the two, or their combination, can contribute differently depending on the environment. In simpler environments like Empty, where the LLM shows good understanding of the environment and can make accurate suggestions, policy influencing is capable of contributing the most. However, as environment complexity increases and the LLM may not be as precise in policy influencing, it seems to pay off to prioritize the more indirect method of reward shaping and having policy influencing contribute less.
